---
title: "Random Walk Simulations"
author: "Jeff Houlahan"
date: "3/4/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Single time series function called popts

This function creates a single simulated population time series
based on a random walk.   
Step 1: initialise an empty vector   
Step 2: Set the initial value by choosing from a uniform distribution between a specified 
  minimum and maximum.   
  Step 3: Create a for loop that adds some normally distributed random noise to N~t~
to create N~t-1~.   


```{r}
popts<-function(n, mn, mx,u,v,w){
m <- vector("numeric", n)

m[1]<-runif(1,min = mn,max = mx)

for (i in 1:(n-1)) {
  
  m[i+1]<-m[i] + (rnorm(u,v, w))
  
}
exp(m)
}
```

## Creating and indexing multiple time series using the popts function called Multipopts

This function creates many simulated time series using popts and indexes each of the time series.   

Step 1: Initialize an empty list.   
Step 2: Create a matric with 2 columns and nrows = the number of years in the time series.
Column 1 holds the index and Column 2 holds the population time series.   
Step 3: Create a for loop that runs the popts as many time as necessary to get the number of time 
series required.   
Step 4: Store each time series in the list.   
Step 5: Bind all the time series into a single data frame.   

```{r}
Multipopts<-function(n, z,mn,mx,u,v,w){
g<-list()
t<-matrix(nrow=n, ncol=2)
for (j in 1:z) {

 t[,2]<-popts(n=n,mn=mn,mx=mx,u=u,v=v,w=w)
 t[,1]<-j
  g[[j]]<-t
}
big_data = do.call(rbind, g)
big_data
}
```

## Example   

Run an example where we want 750 random walk time series of 25 years each with a starting value between 80 and 100 and steps of 1 sd.   
Input would look like Multipopts(n=25,z=750, mn=80, mx=100, u=1, v=0, w=1)

```{r}
Multipopts(n=25,z=750, mn=4, mx=4.5, u=1, v=0, w=1)
```

